---
title: "ESM204_a4_Art_Kristin"
author: "K.H.Art"
date: "2023-05-23"
output: html_document
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(dplyr)
library(purrr)
library(kableExtra)
library(ggforce)
library(cowplot)
library(patchwork)
```
### Calculating the social cost of carbon (SCC) and policy choice under uncertainty
```{r}
#load in data 
damages_raw = read_csv(here("a4/damages.csv")) 
warming_raw = read_csv(here("a4/warming.csv")) 
```

# {.tabset}


## Question 1
Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Omit an intercept term; damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

```{r plot_1, out.width="75%", fig.align = 'center', fig.cap = "**Figure 1:** Estimated annual total damages from climate change at different levels of global mean temperature warming."}
#exploratory plot
plot_0 = ggplot(damages_raw, aes(x = warming, y = damages)) +
  geom_point(shape = 1, color = 'coral3')+
  labs(x = "Warming (◦C)", y = "Damages ($)")+
  theme_minimal()

#plot_0

#add a warming squared value to df
damages_df = damages_raw %>% 
  mutate(warming2 = (warming)^2)

#fit a quadratic regression model to damages data
qm = lm(damages ~ 0 + warming + warming2, data = damages_df)
summary(qm)

#add qm to plot using geom_smooth (option 1)
plot_1 = plot_0 +
  geom_smooth(method = "lm", formula = y ~ x + I(x^2), se = FALSE, color = 'black')

plot_1

#add qm to plot using predicted values (option 2)
#create sample data to visualize model
warming_samples = as.numeric(seq(from = 0, to = 10))
damages_predict = as.numeric(predict(qm, list(warming = warming_samples, warming2 = warming_samples^2)))

#combine into one df 
damage_predictions_df = cbind(warming_samples, damages_predict) %>% 
  as.data.frame()

#add to plot
plot_1 = plot_0 +
  geom_line(data = damage_predictions_df, aes(x = warming_samples, y = damages_predict))

#plot_1

```


## Question 2
Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate trajectory and under the pulse scenario. Make four plots: (1) damages over time without the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that arises from the pulse, and (4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).
```{r plot_2, out.width="75%", fig.align = 'center', fig.cap = "**Figure 2:** Estimated annual total damages from climate change between 2021-2022 based on different global mean climate warming trajectories. A) Damages from the baseline future climate warming trajectory, B) damages from the climate warming trajectory that includes a one-time, 35 billion ton pulse of carbon to the atmosphere today, C) the difference in damages between the baseline and pulse trajectories, D) the difference in damages between the baseline and pulse trajectories per ton of CO2."}
#define function for damages based on qm
compute_damages = function(warming){
  #damage = 0 + a*warming + b*warming^2
  damage = 0 + (qm$coefficients['warming'])*warming + (qm$coefficients['warming2'])*(warming^2)
  
  return(damage)
}

#apply damages function and calculate the things
warming_df = warming_raw %>% 
  mutate(est_damage_bl = purrr::map_dbl(.x = warming_baseline, .f = compute_damages), #compute damages for baseline scenario
         est_damage_p = purrr::map_dbl(.x = warming_pulse, .f = compute_damages), #compute damages for pulse scenario
         diff_damages = est_damage_p - est_damage_bl, #compute difference between above two 
         diff_per_ton = diff_damages/35000000000, #compute difference per ton
         year = as.factor(year), #make year a factor for plotting
         t = as.numeric(seq(0, 79, 1)))

 #pivot longer
warming_df_longer = warming_df %>% 
  select(-c(warming_baseline, warming_pulse)) %>% #remove columns we don't need
  pivot_longer(cols = "est_damage_bl":"diff_per_ton", names_to = "variable", values_to = "damage") %>% 
  mutate(across(variable, factor, levels = c("est_damage_bl", "est_damage_p", "diff_damages", "diff_per_ton"))) #make variable column a factor with levels to control order of subplots below

#plot estimates
plot_2 = ggplot(warming_df_longer, aes(x = year)) +
  geom_point(aes(y = damage), color = "coral3") +
  facet_wrap(~ variable, scales = "free_y", nrow = 2, labeller = as_labeller(c(est_damage_bl = 'A) Baseline Trajectory', est_damage_p='B) Pulse Trajectory', diff_damages = "C) Difference", diff_per_ton ='D) Difference per ton CO2'))) +
  labs(x = "Year", y = "Damages ($)") +
  theme_minimal()

plot_2
```


## Question 3
The SCC is defined as the present discounted value of the stream of future damages caused by one additional ton of CO2.*1* The Obama Administration used a discount rate of 3% to discount damages. The recent EPA update uses a discount rate that changes over time, but starts at 2%. Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for discount rates ranging from 1% to 8%. *1* For the purposes of this homework, you may ignore damages after the year 2100.
```{r plot_3, out.width="75%", fig.align = 'center', fig.cap = "**Figure 3:** Estimated social cost of carbon (SCC) as a function of discount rate."}
#define function to calculate generic PV
calculate_PV = function(range, initial_damage, r){
  #initialize total value
  total = 0
  
  #run for loop over specified range to calculate PV with discount 
  for (i in range){
    total = total + initial_damage/(1 + r)^i
  }
  return(total)
}

#define discount rates
r = seq(from = 0.01, to = 0.08, by = 0.001)

#apply function to discount rates
SCC = map(.x = r, initial_damage = warming_df$diff_per_ton[warming_df$t == 0], range = warming_df$t, .f = calculate_PV) %>%
  unlist() %>% 
  as.data.frame() %>% 
  rename("SCC" = 1)

SCC_results = cbind(r, SCC)

#plot results
plot_3 = ggplot(SCC_results, aes(x = r, y = SCC))+
  geom_line(size = 1)+
  labs(x = "Discount rate (%)", y = "SCC ($/ton)")+
  theme_minimal()

plot_3

```
Higher discount rates result in a lower SCC measurement. 

## Question 4
The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation, a recommendation the EPA considered. Recall the Ramsey Rule: r = ρ + ηg
Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

 ́ The “Ramsey Rule”: r = η g + δ
 ́ η: elasticity of marginal utility of consumption with respect to income (how much more does a poor person value a dollar compared to a rich person?)
 ́ δ: pure rate of time preference
 ́ g: growth rate of income over time
 
```{r plot_4, out.width="75%", fig.align = 'center', fig.cap = "**Figure 4:** Estimated social cost of carbon (SCC) as a function of discount rate. Red dot represents the discount rate and SSC calculated by the Ramsey Rule, which was recommended by the National Academies of Sciences, Engineering, and Medicine."}
#define variables
q = 0.001
n = 2
g = 0.01

#calculate discount rate and resulting SSC
rr = as.numeric((q + n*g))
rr_SSC = as.numeric(calculate_PV(rr, initial_damage = warming_df$diff_per_ton[warming_df$t == 0], range = warming_df$t))
 
#add to plot
plot_4 = plot_3 +
  geom_point(aes(x = rr, y = rr_SSC), color = 'coral3', size = 4) +
  geom_text(aes(x = rr, y = rr_SSC), label = "(0.02, 23.50)", hjust = -0.2, vjust = 0.2)

plot_4
```


## Question 5
Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual emissions, without any substantial climate change mitigation. Policy B is to take immediate and strong action on climate change. Use these facts:
• If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).
• Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever.
• Society is risk neutral
• Use a discount rate of 2%
What is the expected present value of damages up to 2100 under Policy A? What is the
expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy
A costs zero and undertaking Policy B costs X. How large could X be for it to still make economic sense to pursue Policy B instead of Policy A? Qualitatively, how would your answer change if society were risk averse?

- Policy A: Warming will either occur according to the baseline dataset at a probability of 0.5 or warming each year will be 1.5 times higher than that in the baseline dataset at a probability of 0.5

```{r}
#define discount rate of 2%
r = 0.02

#define policy A variables
p = 0.5
coeff = 1.5
c = 0

#define function to calculate expected present value under policy A
compute_EPV_A = function(r, coeff, c, p){
  #calculate the total cost of baseline scenario
  total_bl = as.numeric(calculate_PV(r = r, initial_damage = warming_df$est_damage_bl[warming_df$t == 0], range = warming_df$t))
  
  #calculate the total cost of 1.5 times baseline scenario
  total_1.5 = as.numeric(coeff*total_bl)
  
  #add total benefits, subtract total costs, and multiply by probability 
  EPV_A = (total_bl + total_1.5 - c)*p
  return(EPV_A)
}

EPV_A = compute_EPV_A(r = r, coeff = coeff, c = c, p = p)

EPV_A

```

- Policy B: Warming will continue until 2050 according to the baseline dataset and then stabilize at 1.29 degrees and stay that way forever. 

```{r}
#policy B: take immediate and strong action against climate change
#define policy B variables
p = 1
c = 0

compute_EPV_B = function(r, c, p){
  
  #calculate the total cost of baseline scenario until year 2050
  total_pre_2050 = as.numeric(calculate_PV(r = r, initial_damage = warming_df$est_damage_bl[warming_df$t == 0], range = seq(0, 29, 1)))
  
  #calculate the total cost once temp stabilizes in 2050 
  total_post_2050 = as.numeric(calculate_PV(r = r, initial_damage = warming_df$est_damage_bl[warming_df$year == 2050], range = seq(0, 50, 1)))
  
  #add the total benefits, subtract total costs and multiply by probability
  EPV_B = (total_pre_2050 + total_post_2050 - c)*p
  
  return(EPV_B)
}

#cost of implementation is not included yet
EPV_B = compute_EPV_B(r = r, c = c, p = p)
EPV_B

#calculate difference in damages
EPV_diff = EPV_B - EPV_A
EPV_diff

#now include implementation cost of Policy B 
c = EPV_diff
EPV_B2 = compute_EPV_B(r = r, c = c, p = p)

#check that EPV_B with implementation cost included is equal to EPV_A with implementation cost included
(round(EPV_B2) == round(EPV_A)) #yes, true
```

```{r}
EPV_results = cbind(EPV_A, EPV_B) %>% 
  as.data.frame()

table_1 = kable(EPV_results, format = "markdown", caption = "Table 1: Expected Present Value of damages up to 2100 under Policy A and Policy B.")

table_1
```

The expected present value (EPV) of damages up to 2100 under Policy A is `r EPV_A` while the expected present value (EPV) of damages up to 2100 under Policy B is `r EPV_B`. Policy B could cost up to `r EPV_diff` to implement and still make economic sense to pursue over Policy A if Policy A cost $0 to implement and the society is risk-neutral. If society were risk averse, Policy B would be more attractive because the outcome cost is certain (due to the 100% probability). In Policy A however, the outcome cost is uncertain. There is a 50% chance warming will follow the baseline and result in lower costs, but there's also a 50% risk that warming will be 1.5 times higher than the baseline and cause more damage.

 New concept: “expected value”
 ́ Suppose there are N possible “states of the world”
 ́ Let Bn and Cn be benefit and cost under state of the world n.  ́ Let pn be the probability of state n
Expected Net Benefit (expected value) = p1(B1-C1) + p2(B2-C2)+...+ pN(BN-CN)

```{r plot_7, out.width="75%", fig.align = 'center', fig.cap = "**Figure 5:** Expected Utility of Policy Options for Risk Neutral and Risk Averse Societies."}
#graph results based on risk neutral society
x = seq(0, 5e15, 1e15)
y = seq(0, 5, 1)
#y = seq(0, 14, 1)

df = cbind(x,y) %>% 
  as.data.frame()

plot_5 = ggplot(df, aes(x = x, y = y))+
  geom_line()+
  geom_point(aes(x = EPV_A, y = EPV_A/1e15), color = 'coral3', size = 4) +
  geom_point(aes(x = EPV_B, y = EPV_B/1e15), color = 'coral3', size = 4) +
  geom_text(aes(x = EPV_A, y = EPV_A/1e15), label = "(EPV(A), EU(A))", hjust = -0.2, vjust = 0.2)+
  geom_text(aes(x = EPV_B, y = EPV_B/1e15), label = "(EPV(B), EU(B))", hjust = -0.2, vjust = 0.2)+
  labs(x = "EPV ($)", y = "EU", title = "Expected Utility of Policy Options for Risk Neutral Utility Curve")+
  theme_minimal()

#plot_5

#define square root function for risk averse
sqrt_fun <- function(x) {
  y = (sqrt(x/1e14))
}

#plot risk averse
plot_6 = ggplot(df, aes(x = x, y = y)) +
  geom_function(fun = sqrt_fun)+
  geom_point(aes(x = EPV_A, y = sqrt_fun(EPV_A)), color = 'coral3', size = 4) +
  geom_point(aes(x = EPV_B, y = sqrt_fun(EPV_B)), color = 'coral3', size = 4) +
  geom_text(aes(x = EPV_A, y = sqrt_fun(EPV_A)), label = "(EPV(A), EU(A))", hjust = -0.2, vjust = 0.2)+
  geom_text(aes(x = EPV_B, y = sqrt_fun(EPV_B)), label = "(EPV(B), EU(B))", hjust = -0.2, vjust = 0.2)+
  labs(x = "EPV ($)", y = "EU", title = "Expected Utility of Policy Options for Risk Averse Utility Curve")+
  theme_minimal()

#plot_6

plot_7 = plot_5 + plot_6
plot_7

```


#NOTES

```{r}
# #define function to calculate SSC
# calculate_SCC = function(initial_damage, r, t, row_n){
#   total = 0
#   
#   for (i in warming_df$t){
#     total = total + initial_damage/(1 + r)^i
#   }
#   return(TC)
# }
# 
# #define discount rates
# r = seq(from = 1, to = 8, by = 0.25)
# 
# #apply function to discount rates
# SCC = map(.x = r, initial_damage = warming_df$est_damage_bl[warming_df$t == 0], row_n = nrow(warming_df), .f = calculate_SCC) %>%
#   unlist() %>% 
#   as.data.frame() %>% 
#   rename("SCC" = 1)
# 
# SCC_results = cbind(r, SCC)
# 
# #plot results
# plot_3 = ggplot(SCC_results, aes(x = r, y = SCC))+
#   geom_line(size = 1)+
#   labs(x = "Discount rate (%)", y = "SCC ($/ton)")+
#   theme_minimal()
# 
# plot_3
```