---
title: "ESM204_a2"
author: "Kristin Art & Simone Berkovitz"
date: "2023-04-26"
output: html_document
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(dplyr)
library(ggpubr) 
library(patchwork)
library(cowplot)
library(equatiomatic)
library(pracma)
library(purrr)
library(wesanderson)
```

## Simulating Carbon Abatement

In this assignment, we simulate alternative approaches for reducing carbon emissions among four UC campuses (UCSB, UCLA, UCI, and UCD). We begin with a data set that contains estimates of the marginal cost of reducing carbon emission for each campus for different levels of abatement (in tons of carbon). Note that current (i.e. baseline) emission for each campus are: 

- UCSB = 90,000 tons

- UCLA = 250,000 tons

- UCI = 100,000 tons

- UCD = 160,000 tons

```{r}
#import data and pivot longer
emissions = read_csv(here::here("a2","CarbonUC.csv")) %>% 
  janitor::clean_names() %>% 
  pivot_longer(cols = ucsb:ucd, names_to = "site", values_to = "mac")

#define baseline emissions for each campus (all in tons)
ucsb_baseline = 90000
ucla_baseline = 250000
uci_baseline = 100000
ucd_baseline = 160000

```

### Part 1.  Marginal Abatement Cost Curves
First, we plotted the marginal abatement cost (MAC) against the abatement, or emissions reduction, data from the initial data set. Then, we estimated a linear regression model with a zero intercept for each campus; the resulting marginal abatement cost curve equation for each campus is shown below. 

```{r}
#plot marginal abatement curves
plot_1 = 
  ggplot(emissions, aes(x = abatement, y = mac, color = site)) + 
  geom_jitter() + 
  geom_smooth() + 
  #geom_abline(slope = ucsb_slope)+
  #scale_colour_manual(values = c("#679289","#9C964A","#FDD262","#D6604D")) +
  scale_color_manual(values= wes_palette("GrandBudapest1", n = 4)) +
  labs(title = "Marginal Abatement Cost Curves for UC Campuses", y = "Marginal Abatement Cost ($/ton of carbon)", x = "Abatement (tons of carbon)", color = "Campus") +
  theme_minimal() + 
  ylim(0,700) 

plot_1
  
#estimate linear regression models with a forced 0 intercept, and extract/save model slopes
ucsb_lm = lm(mac ~ 0 + abatement, data = emissions %>% filter(site == "ucsb")) 
ucsb_slope = ucsb_lm$coefficient %>% as.numeric()

ucla_lm = lm(mac ~ 0 + abatement, data = emissions %>% filter(site == "ucla")) 
ucla_slope = ucla_lm$coefficient %>% as.numeric()

uci_lm = lm(mac ~ 0 + abatement, data = emissions %>% filter(site == "uci")) 
uci_slope = uci_lm$coefficient %>% as.numeric()

ucd_lm = lm(mac ~ 0 + abatement, data = emissions %>% filter(site == "ucd")) 
ucd_slope = ucd_lm$coefficient %>% as.numeric()

#add baselines and slopes to full df
emissions_mod = emissions %>%   
  mutate(
    baseline = case_when(
      site == 'ucsb' ~ ucsb_baseline,
      site == 'ucla' ~ ucla_baseline,
      site == 'uci' ~ uci_baseline,
      site == 'ucd' ~ ucd_baseline),
    slope = case_when(
      site == 'ucsb' ~ ucsb_slope,
      site == 'ucla' ~ ucla_slope,
      site == 'uci' ~ uci_slope,
      site == 'ucd' ~ ucd_slope))

#write out the MAC curve equations
#UCSB MAC:
equatiomatic::extract_eq(ucsb_lm, use_coefs = TRUE, coef_digits = 8)
#UCLA MAC:
equatiomatic::extract_eq(ucla_lm, use_coefs = TRUE, coef_digits = 8)
#UCD MAC:
equatiomatic::extract_eq(ucd_lm, use_coefs = TRUE, coef_digits = 8)
#UCI MAC:
equatiomatic::extract_eq(uci_lm, use_coefs = TRUE, coef_digits = 8)
```

### Part 2. Demand Curves
Next, we derived each campus' demand curve for carbon using the estimated MAC curves and known baseline emissions. We plotted those curves and saw that UCD is willing to pay the most for the first ton of carbon emissions.

```{r}
#define generic demand function to solve for P
compute_demand_P = function(slope, baseline, emissions){
  demand = slope*(baseline - emissions)
  return(demand)
}

#apply demand function and save results in main df
emissions_mod = emissions_mod %>% 
  mutate(demand = compute_demand_P(slope, baseline, abatement))

#create linear models of demand (not necessary)
ucsb_d_df = emissions_mod %>% 
  filter(site == "ucsb")
ucsb_d_lm <- lm(demand ~ abatement, data = ucsb_d_df)

ucla_d_df = emissions_mod %>% 
  filter(site == "ucla")
ucla_d_lm <- lm(demand ~ abatement, data = ucla_d_df)

uci_d_df = emissions_mod %>% 
  filter(site == "uci")
uci_d_lm <- lm(demand ~ abatement, data = uci_d_df)

ucd_d_df = emissions_mod %>% 
  filter(site == "ucd")
ucd_d_lm <- lm(demand ~ abatement, data = ucd_d_df)

#write out the demand equations
#UCSB demand curve
equatiomatic::extract_eq(ucsb_d_lm, use_coefs = TRUE, coef_digits = 8)
#UCLA demand curve
equatiomatic::extract_eq(ucla_d_lm, use_coefs = TRUE, coef_digits = 8)
#UCD demand curve
equatiomatic::extract_eq(ucd_d_lm, use_coefs = TRUE, coef_digits = 8)
#UCI demand curve
equatiomatic::extract_eq(uci_d_lm, use_coefs = TRUE, coef_digits = 8)

#plot demand curve for all sites
plot_2 = 
  ggplot(emissions_mod, aes(x = abatement, y = demand, color = site)) +
  geom_line(size = 1) +
  #scale_colour_manual(values = c("#679289","#9C964A","#FDD262","#D6604D")) +
  scale_color_manual(values= wes_palette("GrandBudapest1", n = 4)) +
  labs(x = "Pollution (tons of carbon)", y =  "Demand ($/ ton of carbon )", color = "Campus", title = "Demand Curves for UC Campuses") +
  theme_minimal() +
  theme()+
  ylim(0,500)

plot_2
```

### Part 3. Interventions
Next, we analyzed four alternative interventions for reducing carbon emissions across these four campuses. 

#### a. Intervention #1: Carbon Tax
Under this intervention, each campus must pay a tax of $50/ton for each ton of emissions. 

Under this policy, each campus would choose to emit the following quantities of carbon:

- UCSB: 40,209 tons CO2

- UCLA: 178,718 tons CO2

- UCD: 143,466 tons CO2

- UCI: 75,096 tons CO2

Under this policy, each campus would choose to abate the following quantities of carbon:

- UCSB: 49,491 tons CO2

- UCLA: 71,282 tons CO2

- UCD: 16,534 tons CO2

- UCI: 24,904 tons CO2

The total amount of emissions across all four campuses is 437,489.3 tons of CO2 and the total tax revenue generated is $21,874,467.

```{r}
#define generic function to calculate abatement Q
compute_demand_Q = function(slope, baseline, price){
  Q = baseline - ((price)/(slope))
  return(Q)
}

#define generic function to calculate abatement Q
compute_abatement_Q = function(baseline, emit_Q){
  abatement_Q = baseline - emit_Q
  return(abatement_Q)
}

#calculate the quantity emitted and abated by each campus
ucsb_emit_Q = compute_demand_Q(ucsb_slope, ucsb_baseline, price = 50)
ucsb_abate_Q = compute_abatement_Q(ucsb_baseline, ucsb_emit_Q)
print(paste0("Under intervention #1 (carbon tax), UCSB will emit " , ucsb_emit_Q, " tons of carbon and abate ", ucsb_abate_Q, " tons of carbon."))

ucla_emit_Q = compute_demand_Q(ucla_slope, ucla_baseline, price = 50)
ucla_abate_Q = compute_abatement_Q(ucla_baseline, ucla_emit_Q)
print(paste0("Under intervention #1 (carbon tax), UCLA will emit " , ucla_emit_Q, " tons of carbon and abate ", ucla_abate_Q, " tons of carbon."))

ucd_emit_Q = compute_demand_Q(ucd_slope, ucd_baseline, price = 50)
ucd_abate_Q = compute_abatement_Q(ucd_baseline, ucd_emit_Q)
print(paste0("Under intervention #1 (carbon tax), UCD will emit " , ucd_emit_Q, " tons of carbon and abate ", ucd_abate_Q, " tons of carbon."))

uci_emit_Q = compute_demand_Q(uci_slope, uci_baseline, price = 50)
uci_abate_Q = compute_abatement_Q(uci_baseline, uci_emit_Q)
print(paste0("Under intervention #1 (carbon tax), UCI will emit " , uci_emit_Q, " tons of carbon and abate ", uci_abate_Q, " tons of carbon."))

#calculate the total quantity emitted and total tax revenue generated
total_emitted = sum(ucsb_emit_Q, ucla_emit_Q, uci_emit_Q, ucd_emit_Q)
total_revenue = total_emitted*50
print(paste0("Under intervention #1 (carbon tax), the four campuses will emit a total of " , total_emitted, " tons of carbon and generate $", total_revenue, " in tax revenue."))

```

#### b. Intervention #2: Carbon Ban
Under this intervention, carbon emissions are banned at each campus. The total cost of this ban to each campus would be as follows:

- UCSB: $4,067,016

- UCLA: $21,919,836

- UCD: $38,709,018

- UCI: $10,038,644

```{r}
#define individual supply functions to solve for P
ucsb_mac_P = function(Q){
  P = ucsb_slope*Q
  return(P)}
ucla_mac_P = function(Q){
  P = ucla_slope*Q
  return(P)}
ucd_mac_P = function(Q){
  P = ucd_slope*Q
  return(P)}
uci_mac_P = function(Q){
  P = uci_slope*Q
  return(P)}

#calculate the cost of the ban for each campus
ucsb_ban_P = abs(integral(f = ucsb_mac_P, xmin = ucsb_baseline, xmax = 0))
ucla_ban_P = abs(integral(f = ucla_mac_P, xmin = ucla_baseline, xmax = 0))
ucd_ban_P = abs(integral(f = ucd_mac_P, xmin = ucd_baseline, xmax = 0))
uci_ban_P = abs(integral(f = uci_mac_P, xmin = uci_baseline, xmax = 0))

#calculate the total cost for all 4 campuses
total_ban_P = sum(ucsb_ban_P, ucla_ban_P, ucd_ban_P, uci_ban_P)

print(paste0("Intervention #2 (carbon ban) will cost $", ucsb_ban_P," at UCSB, $", ucla_ban_P, " at UCLA, $", ucd_ban_P, " at UCD, and $", uci_ban_P, "at UCI. The total cost to all four campuses is $", total_ban_P, "."))

```

#### c. Intervention #3: Cap without Trade
Under this intervention, each campus is restricted to an emissions cap of 100,000 tons of carbon. Each campus would need to abate the following quantities of carbon to comply with this cap:

- UCSB: 0 tons CO2

- UCLA: 150,000 tons CO2

- UCD: 60,000 tons CO2

- UCI: 0 tons CO2

The marginal cost of the last unit of abatement for each campus to comply is:

- UCSB: $0

- UCLA: $105

- UCD: $181

- UCI: $0

The total cost of abatement for each campus to comply is:

- UCSB: $0

- UCLA: $18,412,662

- UCD: $23,588,308

- UCI: $0

```{r}
#apply generic abatement function from above to calculate abatement Q 
cap_Q = 100000

#calculate quantity abated for each campus
ucsb_abate_Q2 = max(0,compute_abatement_Q(ucsb_baseline, cap_Q))
ucla_abate_Q2 = max(0,compute_abatement_Q(ucla_baseline, cap_Q))
ucd_abate_Q2 = max(0,compute_abatement_Q(ucd_baseline, cap_Q))
uci_abate_Q2 = max(0,compute_abatement_Q(uci_baseline, cap_Q))

print(paste0("Under intervention #3 (cap without trade), UCSB will abate ", ucsb_abate_Q2, ", UCLA will abate ", ucla_abate_Q2, ", UCD will abate ", ucd_abate_Q2, ", and UCI will abate ", uci_abate_Q2,"."))

#calculate marginal cost of the last unit of abatement for each campus
ucsb_cap_P = ucsb_mac_P(ucsb_abate_Q2)
ucla_cap_P = ucla_mac_P(ucla_abate_Q2)
ucd_cap_P = ucd_mac_P(ucd_abate_Q2)
uci_cap_P = uci_mac_P(uci_abate_Q2 )

print(paste0("Under intervention #3 (cap without trade), the cost of the last unit of abatement for each campus will be $", ucsb_cap_P, " for UCSB, $", ucla_cap_P, " for UCLA, $", ucd_cap_P, " for UCD, and $", uci_cap_P," for UCD."))

#calculate total cost of abatement for each campus 
ucsb_cap_P2 = max(0,(integral(f = ucsb_mac_P, xmin = cap_Q, xmax = ucsb_baseline)))
ucla_cap_P2 = max(0,(integral(f = ucla_mac_P, xmin = cap_Q, xmax = ucla_baseline)))
ucd_cap_P2 = max(0,(integral(f = ucd_mac_P, xmin = cap_Q, xmax = ucd_baseline)))
uci_cap_P2 = max(0,(integral(f = uci_mac_P, xmin = cap_Q, xmax = uci_baseline)))

print(paste0("Under intervention #3 (cap without trade), the total cost for each campus to comply with the cap is $", ucsb_cap_P2, " for UCSB, $", ucla_cap_P2, " for UCLA, $", ucd_cap_P2, " for UCD, and $", uci_cap_P2," for UCD."))

```

#### d. Intervention #4: Cap with Trade
Under this intervention, each campus is restricted to an emissions cap of 100,000 tons of carbon, but trade is allowed. With trade, total emissions will be 400,000 tons of carbon, but some campuses may emit more or less than 100,000. Under this intervention, the equilibrium trading price of carbon permits after trade would be $61.53/ ton. After trade, each campus would emit the following quantities of carbon:

- UCSB: 28,723 tons CO2

- UCLA: 162,273 tons CO2

- UCD: 139,652 tons CO2

- UCI: 69,351 tons CO2

UCSB and UCI would be sellers of carbon permits while UCLA and UCD would be buyers of carbon permits under this scenario.

```{r}
#define individual demand functions to solve for Q
ucsb_d_Q = function(P){
  Q = (((ucsb_slope*ucsb_baseline) - P)/ ucsb_slope)
  return(Q)
}

ucla_d_Q = function(P){
  Q = (((ucla_slope*ucla_baseline) - P)/ ucla_slope)
  return(Q)
}

ucd_d_Q = function(P){
  Q = (((ucd_slope*ucd_baseline) - P)/ ucd_slope)
  return(Q)
}

uci_d_Q = function(P){
  Q = (((uci_slope*uci_baseline) - P)/ uci_slope)
  return(Q)
}

#define function to calculate aggregate demand curve and solve for Q
agg_d_Q = function(P){
  Q = (((ucsb_slope*ucsb_baseline) - P)/ ucsb_slope) + (((ucla_slope*ucla_baseline) - P)/ ucla_slope) + (((ucd_slope*ucd_baseline) - P)/ ucd_slope) + (((uci_slope*uci_baseline) - P)/ uci_slope)
  return(Q)
}

#find total baseline emissions and slope for all 4 campuses to rearrange aggregate demand curve below
total_baseline = sum(ucsb_baseline, ucla_baseline, ucd_baseline, uci_baseline)
total_slope = ((1/ucsb_slope) + (1/ucla_slope) +(1/uci_slope)+(1/ucd_slope))

#define function to calculate aggregate demand curve and solve for P
agg_d_P = function(Q){
  P = ((total_baseline - Q)/total_slope)
  return(P)
}

#calculate the equilibrium trading price
cap_trade_Q = 400000
cap_trade_P = agg_d_P(cap_trade_Q)

#calculate quantity emitted by each campus after trade and compare to cap value ($100,000) to determine who is a seller and who is a buyer
ucsb_cap_Q = ucsb_d_Q(cap_trade_P)
ucla_cap_Q = ucla_d_Q(cap_trade_P)
ucd_cap_Q = ucd_d_Q(cap_trade_P)
uci_cap_Q = uci_d_Q(cap_trade_P)

#check that this is equal to the total cap, 400,000
total_Q = sum(ucsb_cap_Q, ucla_cap_Q, ucd_cap_Q, uci_cap_Q)
print(total_Q == cap_trade_Q)

print(paste0("Under intervention #4 (cap with trade) where the total cap is set at 400,000, the equilibrium trading price of permits after trade would be ", cap_trade_P, ". Under this scenario, UCSB will emit ", ucsb_cap_Q, " tons of carbon, UCLA will emit ", ucla_cap_Q, " tons of carbon, UCD will emit ", ucd_cap_Q, " tons of carbon, and UCI will emit ", uci_cap_Q," tons of carbon. UCSB and UCI are sellers of carbon emission permits while UCLA and UCD are buyers of carbon emissions permits in this market."))

```

### Part 4. Philanthropy and Cap & Trade
Next, we analyzed what would happen if the central campus decides to go with the "Cap with Trade" approach (Intervention #4 above) and an altruistic philanthropist buys and retires a portion of the available permits.
That philanthropist would pay a total of $3,076,721 to purchase and retire 50,000 permits at the equilibrium trading price from above. 
With the reduced cap of 350,000 permits, the new trading price among the four campuses would be $76.92.
```{r}
#calculate the new trading price for the 4 campuses
cap_trade_Q2 = 350000
cap_trade_P2 = agg_d_P(cap_trade_Q2)
cap_trade_P2

#calculate the cost for the philanthropist How much will it cost the philanthropist to buy and retire these permits?
phil_cost = 50000*cap_trade_P
phil_cost

print(paste0("A philanthropist will have to pay $", phil_cost, " to purchase 50,000 of the permits under intervention #4 (cap with trade). The new trading price for the 4 campuses will be ", cap_trade_P2, " with the reduced total emissions cap of 350,000 tons of carbon."))


```
